---
title: "Homework 2"
author: "Karthika Shankar"
date: "2026-10-07"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

# Problem 1
```{r, warning=FALSE}
subway_df = 
  read_csv(
    
  "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(
      Route8 = col_character(),
      Route9 = col_character(),
      Route10 = col_character(),
      Route11 = col_character()
    )
  ) |>
  janitor::clean_names() |>
  select(
    line,
    station_name,
    station_latitude,
    station_longitude,
    route1:route11,
    entry,
    vending,
    entrance_type,
    ada
  ) |>
  mutate(
    entry = case_match(
      entry,
      "YES" ~ TRUE,
      "NO" ~ FALSE
    )
  )
```

The subway dataset includes information about subway stations and their entrances and exits like the station name, line, location, routes served, entrance type, vending availability, and ADA accessibility. I cleaned the variable names, kept only the variables needed for this problem, and changed the `entry` variable from "YES" and "NO" to a logical variable. The cleaned dataset has `r nrow(subway_df)` rows and `r ncol(subway_df)` columns. The data are not tidy because the routes are spread across many different columns (`route1` through `route11`) instead of being stored in one route variable.

```{r}
subway_df |>
  distinct(line, station_name) |>
  nrow()
```

There are 465 distinct subway stations in the dataset.

```{r}
subway_df |>
  filter(ada == TRUE) |>
  distinct(line, station_name) |>
  nrow()
```

There are 84 ADA-compliant subway stations in the dataset.

```{r}
subway_df |>
  filter(vending == "NO") |>
  pull(entry) |>
  mean()
```

Among station entrances and exits without vending about 37.7% allow entrance.

```{r}
subway_tidy_df =
  subway_df |>
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name"
  ) |>
  drop_na(route_name)
```

```{r}
subway_tidy_df |>
  filter(route_name == "A") |>
  distinct(line, station_name) |>
  nrow()
```

There are 60 distinct subway stations that serve the A train.

```{r}
subway_tidy_df |>
  filter(route_name == "A" & ada == TRUE) |>
  distinct(line, station_name) |>
  nrow()
```

Of the stations that serve the A train 17 are ADA compliant.

# Problem 2
```{r}
mr_trash_df =
  readxl::read_excel(
    "./data/202610 Trash Wheel Collection Data.xlsx",
    sheet = "Mr. Trash Wheel",
    range = "A2:N742"
  ) |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    year = as.factor(year),
    sports_balls = as.integer(round(sports_balls)),
    trash_wheel = "Mr. Trash Wheel"
  )
```

```{r}
professor_trash_df =
  readxl::read_excel(
    "./data/202610 Trash Wheel Collection Data.xlsx",
    sheet = "Professor Trash Wheel",
    range = "A2:M139",
    na = c("", "dive")
  ) |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    year = as.factor(year),
    trash_wheel = "Professor Trash Wheel"
  )
```

```{r}
gwynnda_df =
  readxl::read_excel(
    "./data/202610 Trash Wheel Collection Data.xlsx",
    sheet = "Gwynnda the Good Wheel of the W",
    range = "A2:L394"
  ) |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    year = as.factor(year),
    trash_wheel = "Gwynnda"
  )
```

```{r}
trash_wheel_df =
  bind_rows(
    mr_trash_df,
    professor_trash_df,
    gwynnda_df
  )
```

```{r}
professor_weight =
  trash_wheel_df |>
  filter(trash_wheel == "Professor Trash Wheel") |>
  pull(weight_tons) |>
  sum(na.rm = TRUE)

gwynnda_cigarettes =
  trash_wheel_df |>
  filter(
    trash_wheel == "Gwynnda",
    month == "June",
    year == "2022"
  ) |>
  pull(cigarette_butts) |>
  sum(na.rm = TRUE)
```

```{r}
nrow(trash_wheel_df)
professor_weight
gwynnda_cigarettes
```

The combined Trash Wheel dataset has `r nrow(trash_wheel_df)` observations from Mr. Trash Wheel, Professor Trash Wheel, and Gwynnda. The dataset includes information like the dumpster number, month and year, weight and volume of trash collected, cigarette butts, plastic bottles, and other types of waste. I added a `trash_wheel` variable so it is easy to see which wheel each observation came from. Professor Trash Wheel collected `r professor_weight` tons of trash in total, and Gwynnda collected `r format(gwynnda_cigarettes, big.mark = ",", scientific = FALSE)` cigarette butts in June 2022.


# Problem 3
```{r}
mci_baseline_df =
  read_csv(
    "./data/MCI_baseline.csv",
    skip = 1,
    na = "."
  ) |>
  janitor::clean_names()
```

```{r}
mci_baseline_df =
  mci_baseline_df |>
  mutate(
    sex =
      case_match(
        sex,
        1 ~ "male",
        0 ~ "female"
      ),
    sex = as.factor(sex),
    apoe4 =
      case_match(
        apoe4,
        1 ~ "carrier",
        0 ~ "non-carrier"
      ),
    apoe4 = as.factor(apoe4)
  ) |>
  filter(
    is.na(age_at_onset) | age_at_onset > current_age
  )
```

```{r}
number_recruited =
  nrow(mci_baseline_df)

number_developed_mci =
  mci_baseline_df |>
  drop_na(age_at_onset) |>
  nrow()

average_baseline_age =
  mci_baseline_df |>
  pull(current_age) |>
  mean()

women_apoe4 =
  mci_baseline_df |>
  filter(sex == "female") |>
  mutate(apoe4_carrier = apoe4 == "carrier") |>
  pull(apoe4_carrier) |>
  mean()
```

```{r}
number_recruited
number_developed_mci
average_baseline_age
women_apoe4
```

The baseline dataset originally had 483 participants and included information on age, sex, education, APOE4 status, and age at MCI onset. I skipped the first descriptive row and treated "." as a missing value when importing the data. I changed sex and APOE4 status into categorical variables and removed participants who already had MCI at baseline. After cleaning, `r number_recruited` participants were left and `r number_developed_mci` of them developed MCI during follow-up. The average baseline age was `r round(average_baseline_age, 2)` years and `r round(women_apoe4 * 100, 1)`% of the women were APOE4 carriers.

```{r}
mci_amyloid_df =
  read_csv(
    "./data/mci_amyloid.csv",
    skip = 1,
    na = c(".", "NA", "", "Na")
  ) |>
  janitor::clean_names() |>
  rename(id = study_id) |>
  pivot_longer(
    baseline:time_8,
    names_to = "visit",
    values_to = "amyloid"
  )
```

The amyloid dataset includes measurements from baseline and several follow-up visits. I skipped the first descriptive row, cleaned the variable names, handled the missing values, and used `pivot_longer()` to organize the repeated amyloid measurements.

```{r}
baseline_only =
  mci_baseline_df |>
  distinct(id) |>
  left_join(
    mci_amyloid_df |>
      distinct(id) |>
      mutate(in_amyloid = TRUE),
    by = "id"
  ) |>
  filter(is.na(in_amyloid))

amyloid_only =
  mci_amyloid_df |>
  distinct(id) |>
  left_join(
    mci_baseline_df |>
      distinct(id) |>
      mutate(in_baseline = TRUE),
    by = "id"
  ) |>
  filter(is.na(in_baseline))
```

```{r}
nrow(baseline_only)
nrow(amyloid_only)
```

```{r}
mci_df =
  inner_join(
    mci_baseline_df,
    mci_amyloid_df,
    by = "id"
  )
```

```{r}
mci_df |>
  distinct(id) |>
  nrow()

nrow(mci_df)
```

There are 8 participants who appear only in the baseline dataset and 16 who appear only in the amyloid dataset. After combining the datasets and keeping only participants who appear in both, the final dataset has 471 participants and 2355 observations. The combined dataset includes the baseline demographic information with each participant's visit and amyloid measurement.

```{r}
write_csv(
  mci_df,
  "./data/mci_clean.csv"
)
```