---
title: "Homework 2"
author: "Karthika Shankar"
date: "2026-10-07"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

# Problem 1
```{r}
subway_df = 
  read_csv(
    
  "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(
      Route8 = col_character(),
      Route9 = col_character(),
      Route10 = col_character(),
      Route11 = col_character()
    )
  ) |>
  janitor::clean_names() |>
  select(
    line,
    station_name,
    station_latitude,
    station_longitude,
    route1:route11,
    entry,
    vending,
    entrance_type,
    ada
  ) |>
  mutate(
    entry = case_match(
      entry,
      "YES" ~ TRUE,
      "NO" ~ FALSE
    )
  )
```

The subway dataset includes information about subway stations and their entrances and exits like the station name, line, location, routes served, entrance type, vending availability, and ADA accessibility. I cleaned the variable names, kept only the variables needed for this problem, and changed the `entry` variable from "YES" and "NO" to a logical variable. The cleaned dataset has `r nrow(subway_df)` rows and `r ncol(subway_df)` columns. The data are not tidy because the routes are spread across many different columns (`route1` through `route11`) instead of being stored in one route variable.

```{r}
subway_df |>
  distinct(line, station_name) |>
  nrow()
```

There are 465 distinct subway stations in the dataset.

```{r}
subway_df |>
  filter(ada == TRUE) |>
  distinct(line, station_name) |>
  nrow()
```

There are 84 ADA-compliant subway stations in the dataset.

```{r}
subway_df |>
  filter(vending == "NO") |>
  pull(entry) |>
  mean()
```

Among station entrances and exits without vending about 37.7% allow entrance.

```{r}
subway_tidy_df =
  subway_df |>
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name"
  ) |>
  drop_na(route_name)
```

```{r}
subway_tidy_df |>
  filter(route_name == "A") |>
  distinct(line, station_name) |>
  nrow()
```

There are 60 distinct subway stations that serve the A train.

```{r}
subway_tidy_df |>
  filter(route_name == "A" & ada == TRUE) |>
  distinct(line, station_name) |>
  nrow()
```

Of the stations that serve the A train 17 are ADA compliant.
