---
title: "Homework 2"
author: "Karthika Shankar"
date: "2026-10-07"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

# Problem 1
```{r}
subway_df = 
  read_csv(
    
  "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(
      Route8 = col_character(),
      Route9 = col_character(),
      Route10 = col_character(),
      Route11 = col_character()
    )
  ) |>
  janitor::clean_names() |>
  select(
    line,
    station_name,
    station_latitude,
    station_longitude,
    route1:route11,
    entry,
    vending,
    entrance_type,
    ada
  ) |>
  mutate(
    entry = case_match(
      entry,
      "YES" ~ TRUE,
      "NO" ~ FALSE
    )
  )
```

The subway dataset includes information about subway stations and their entrances and exits like the station name, line, location, routes served, entrance type, vending availability, and ADA accessibility. I cleaned the variable names, kept only the variables needed for this problem, and changed the `entry` variable from "YES" and "NO" to a logical variable. The cleaned dataset has `r nrow(subway_df)` rows and `r ncol(subway_df)` columns. The data are not tidy because the routes are spread across many different columns (`route1` through `route11`) instead of being stored in one route variable.

```{r}
subway_df |>
  distinct(line, station_name) |>
  nrow()
```

There are 465 distinct subway stations in the dataset.

```{r}
subway_df |>
  filter(ada == TRUE) |>
  distinct(line, station_name) |>
  nrow()
```

There are 84 ADA-compliant subway stations in the dataset.

```{r}
subway_df |>
  filter(vending == "NO") |>
  pull(entry) |>
  mean()
```

Among station entrances and exits without vending about 37.7% allow entrance.

```{r}
subway_tidy_df =
  subway_df |>
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name"
  ) |>
  drop_na(route_name)
```

```{r}
subway_tidy_df |>
  filter(route_name == "A") |>
  distinct(line, station_name) |>
  nrow()
```

There are 60 distinct subway stations that serve the A train.

```{r}
subway_tidy_df |>
  filter(route_name == "A" & ada == TRUE) |>
  distinct(line, station_name) |>
  nrow()
```

Of the stations that serve the A train 17 are ADA compliant.

#Problem 2

```{r}
mr_trash_df =
  readxl::read_excel(
    "./data/202610 Trash Wheel Collection Data.xlsx",
    sheet = "Mr. Trash Wheel",
    range = "A2:N742"
  ) |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    year = as.factor(year),
    sports_balls = as.integer(round(sports_balls)),
    trash_wheel = "Mr. Trash Wheel"
  )
```

```{r}
professor_trash_df =
  readxl::read_excel(
    "./data/202610 Trash Wheel Collection Data.xlsx",
    sheet = "Professor Trash Wheel",
    range = "A2:M139",
    na = c("", "dive")
  ) |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    year = as.factor(year),
    trash_wheel = "Professor Trash Wheel"
  )
```

```{r}
gwynnda_df =
  readxl::read_excel(
    "./data/202610 Trash Wheel Collection Data.xlsx",
    sheet = "Gwynnda the Good Wheel of the W",
    range = "A2:L394"
  ) |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    year = as.factor(year),
    trash_wheel = "Gwynnda"
  )
```

```{r}
trash_wheel_df =
  bind_rows(
    mr_trash_df,
    professor_trash_df,
    gwynnda_df
  )
```

```{r}
professor_weight =
  trash_wheel_df |>
  filter(trash_wheel == "Professor Trash Wheel") |>
  pull(weight_tons) |>
  sum(na.rm = TRUE)

gwynnda_cigarettes =
  trash_wheel_df |>
  filter(
    trash_wheel == "Gwynnda",
    month == "June",
    year == "2022"
  ) |>
  pull(cigarette_butts) |>
  sum(na.rm = TRUE)
```

```{r}
nrow(trash_wheel_df)
professor_weight
gwynnda_cigarettes
```

The combined Trash Wheel dataset has `r nrow(trash_wheel_df)` observations from Mr. Trash Wheel, Professor Trash Wheel, and Gwynnda. The dataset includes information like the dumpster number, month and year, weight and volume of trash collected, cigarette butts, plastic bottles, and other types of waste. I added a `trash_wheel` variable so it is easy to see which wheel each observation came from. Professor Trash Wheel collected `r professor_weight` tons of trash in total, and Gwynnda collected `r gwynnda_cigarettes` cigarette butts in June 2022.



